Adaptive evolution with aneuploidy and mutation

### 2022-01-24 ms.tex
added analysis results

% TODO cite https://www.biorxiv.org/content/10.1101/2020.10.07.327759v1
% TODO cite Yang, F. et al. The fitness costs and benefits of trisomy of each Candida albicans chromosome. Genetics 218, 1–7 (2021). https://academic.oup.com/genetics/article/218/2/iyab056/6218773
% cite: https://www.cell.com/developmental-cell/fulltext/S1534-5807(21)00562-1?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS1534580721005621%3Fshowall%3Dtrue
% cite: https://www.cell.com/developmental-cell/fulltext/S1534-5807(21)00592-X?_returnURL=https%3A%2F%2Flinkinghub.elsevier.com%2Fretrieve%2Fpii%2FS153458072100592X%3Fshowall%3Dtrue

@@

%%%%%%%% Fig illustation
\begin{figure}[b!]
  %\centering
  \begin{subfigure}[t]{0.4\textwidth}
      \caption{
        \textbf{A. Single-locus model}
      }
      \centering
      \includegraphics[height=1.8in]{../figures/Fig1-A.pdf}      
      \label{fig:model1}
  \end{subfigure}
  %\\
  \begin{subfigure}[t]{0.4\textwidth}
  	  \caption{
        \textbf{B. Multi-locus model}
      }
      \centering
      \includegraphics[height=1.8in]{../figures/Fig1-B.pdf}      
      \label{fig:model2}
  \end{subfigure}
  \caption{
    \textbf{Model illustrations.}
    \textbf{(A)} In the single-locus model, the four genotypes are: euploid wild-type, \euwt; euploid mutant, \eumt; aneuploid wild-type, \anwt; and aneuploid mutant, \anmt.
    Overall there are two possible trajectories from \euwt\; to \eumt.
    \textbf{(B)} In the multi-locus model, each genotype is characterized by its karyotype, \euwt\; or \anwt, and the number of accumulated beneficial mutations, denoted by stars. 
    In both panels arrows denote transitions between genotypes, with transitions rates: $\mu$, beneficial mutation rate; $\delta$, aneuploidy rate.
  }
  \label{fig:models}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Prior distributions.} We used informative prior distributions for $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$, which we estimated from growth curves data from mono-culture growth experiments previously reported by \citet[Figs. 3C, 4A, and S2]{Yona2012}.
We used \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. Briefly, \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed culture, that is, it predicts the results of a competition assay.
From these predictions, relative fitness values can be computed. Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. We sampled 10,000 samples to use as a prior distribution~(\autoref{fig:growth-curves}).
We used growth curves of \euwt\ and \anwt\ in \SI{39}{\celsius} to estimate a prior distribution for $w_{2n+1}$. In lieu of a better prior, we used the same prior for $w_{2n+1^*}$ and $w_{2n^*}$. 
To increase computational efficiency, we also assumed $w(2n*)>w(2n+1*)>w(2n+1)>w(2n)$; running the inference without this assumption produced similar results.

@@

% FIG posterior
%% generated with posterior-plot.ipynb
\begin{figure}[h!]
  \centering
\includegraphics[width=0.9\textwidth]{../figures/posterior.pdf}
  \caption{
  \textbf{Posterior distribution of single-locus model.}
On the diagonal, the inferred posterior distribution of each model parameter. 
Below the diagonal, the inferred joint posterior distribution of pairs of model parameters (dark purple and bright yellow for low and high density, respectively). Red markers and orange lines for the joint MAP estimate (which may differ from the marginal MAP, as the marginal distribution integrates over all other parameters).
} 
  \label{fig:posterior}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%

@@

However, a model without aneuploidy where the aneuploidy rate is fixed at zero, $\delta=0$, cannot explain the experimental observations~(Figure~\ref{fig:2n*-fixation}). The estimated parameter values without aneuploidy are: 
%% generated with MAP-and-hdi.ipynb
mutation rate, $\mu=7.837\substack{+0.078 \\ -0.123}\times10^{-9}$,
fitness of mutant, $w_{2n^*}=1.013\substack{+0.000 \\ -0.000}$. 
We can see that the mutation rate estimate in this model is much lower than in the model with aneuploidy. Higher mutation rate would cause quicker fixation of \eumt\ than in the experimental observations (\autoref{fig:2n*-fixation}). Thus, even disregarding the appearance of trisomic cells in the experiment, the evidence supports the inclusion of aneuploidy in the model.
% TODO WAIC?

% Fig single-locus model with and without aneuploidy
%% generated with with-aneuploidy_vs_no_aneuploidy.ipynb
\begin{figure}[h!]
  \begin{subfigure}{0.5\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/fixation-plot-a.pdf}      
      \label{fig:fit}
  \end{subfigure}
  \begin{subfigure}{0.5\textwidth}
      \centering
      \includegraphics[width=\textwidth]{../figures/fixation-plot-b.pdf}      
      \label{fig:fit-cumulative}
  \end{subfigure}
  \caption{
    \textbf{Single-locus model fit with and without aneuploidy.}
    \textbf{(A)} The distribution of time to fixation of \eumt\ (i.e., adaptation time) in $10,000$ simulations of the single-locus model with aneuploidy (blue; MAP parameters) compared to two models without aneuploidy: a model with the same parameter values except $\delta=0$ (orange), and a model fitted to the data assuming $\delta=0$ (green).
In the experiment by \citet{Yona2012}, one population lost aneuploidy by generation 1,700 and another by generation 2,350 (dashed lines) but not before generation 450. Thus, the blue distribution is a better fit compared to the green, and the yellow histogram has a very poor fit.
The last bin contains all the simulations with time equal or greater than 3,000.
    \textbf{(B)}
	Cumulative distribution of the time to fixation of \eumt\ in 10,000 simulations using the MAP estimate with and without aneuploidy in blue and green, respectively, and corresponding to the blue and green bars in panel A. The MAP likelihood (\autoref{eq:heatstress-likelihood}) is 0.84 and 0.67 with and without aneuploidy, respectively.
     }
  \label{fig:2n*-fixation}
\end{figure}

\begin{figure}[h!]
  \centering
\includegraphics[width=0.8\textwidth]{../figures/dynamics.pdf}
  \caption{
  \textbf{Posterior genotype frequency dynamics for the single-locus model.} 
  The posterior prediction for the frequencies of the four genotypes over time. Each of the 20 curves is the average of 10,000 simulations of the single-locus model using parameters drawn from the posterior distribution.
  }
  \label{fig:ppc-plot}
\end{figure}

@@

\begin{figure}[h]
% Fig Curveball 30deg
% generated with growth_curves.ipynb
\begin{subfigure}{0.5\textwidth}
    \centering
	\includegraphics[width=\textwidth]{../figures/evo39_fitness_30deg.pdf}
    \label{fig:growth-curves-30deg}
\end{subfigure}
% Fig Curveball 39deg
% generated with growth_curves.ipynb
\begin{subfigure}{0.5\textwidth}
    \centering
	\includegraphics[width=\textwidth]{../figures/evo39_fitness_39deg.pdf}
    \label{fig:growth-curves-39deg}
\end{subfigure}
\caption{
    \textbf{Fitness estimation from growth curves.}
    \textbf{(A)} Fitness estimation from growth curves collected at \SI{30}.{\celsius}
    \textbf{(B)} Fitness estimation from growth curves collected at \SI{39}.{\celsius}
    Growth curves previously described in \citet[Figs. 3C, 4A, and S2]{Yona2012}.
	Fitness estimated from growth curves using \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. See \emph{Models and Methods, Prior distributions} for more details.
} 
\label{fig:growth-curves}
\end{figure}
%TODO please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
% TODO add panel labels

% Fig likelihood profile
%% generated by sensitivity-analysis.ipynb
\begin{figure}[p]
  \centering
  \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../figures/sensitivity-A.pdf}      
      \label{fig:sensitivity-mutation}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../figures/sensitivity-B.pdf}      
      \label{fig:sensitivity-aneuploidy}
  \end{subfigure}
  \\
   \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../figures/sensitivity-C.pdf}      
      \label{fig:sensitivity-anwt}
  \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../figures/sensitivity-D.pdf}      
      \label{fig:sensitivity-anmt}
  \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{../figures/sensitivity-E.pdf}      
      \label{fig:sensitivity-eumt}
  \end{subfigure}
  \caption{
    \textbf{Likelihood profiles.} Sensitivity of the approximate likelihood of the single-locus model, $\likelihood(\theta)$, to changing a single parameter while the other parameters remain fixed at their MAP estimates. Dashed vertical line represents the MAP value.
  }

  \label{fig:sensitivity}
  % TODO what defines the x range? is it the prior distribution range?
\end{figure}

% Fig ABC iterations, convergence
%% generated with convergence.ipynb
\begin{figure}[h]

  \begin{subfigure}{1\textwidth}
    \centering
      \includegraphics[width=0.45\textwidth]{../figures/convergence-p1_mr.pdf}      
      \includegraphics[width=0.45\textwidth]{../figures/convergence-p2_tr.pdf} \\     
      \includegraphics[width=0.325\textwidth]{../figures/convergence-p3_w1.pdf}      
      \includegraphics[width=0.325\textwidth]{../figures/convergence-p4_w2.pdf}      
      \includegraphics[width=0.325\textwidth]{../figures/convergence-p5_w3.pdf}      
     \label{fig:convergence-A}
    \end{subfigure}
    \\    
  \begin{subfigure}{1\textwidth}
  	\includegraphics[width=\textwidth]{../figures/ess.pdf}
     \label{fig:convergence-B}
  \end{subfigure}
    \caption{
    \textbf{Single-locus model inference convergence.} 
    The ABC-SMC algorithm was used to infer the parameters of the single-locus model. \textbf{(A-E)} The approximate posterior distributions of model parameters at each iteration of the ABC-SMC algorithm demonstrates convergence, as the posterior did not significantly change after the first iteration, $t=1$.
    \textbf{(F-I)} ABC-SMC measures of convergence. After iteration number 6, the acceptance threshold was $\epsilon=0.13$, the acceptance rate was $0.018$, the number of particles was 982, and the effective sample size ESS=651.
}
    \label{fig:convergence}
\end{figure}

% Fig Validation of posteriors
%% generated with diff-runs.ipynb
\begin{figure}[h]
    \centering
      \includegraphics[width=0.45\textwidth]{../figures/runs-A.pdf}      
      \includegraphics[width=0.45\textwidth]{../figures/runs-B.pdf}    
      \includegraphics[width=0.325\textwidth]{../figures/runs-C.pdf}      
      \includegraphics[width=0.325\textwidth]{../figures/runs-D.pdf}      
      \includegraphics[width=0.325\textwidth]{../figures/runs-E.pdf} 
       \caption{
    \textbf{Posterior distribution validation.}
    The posterior distribution of the single-locus model parameters is roughly the same regardless of the number of simulations (4-10,000 replicates) used to approximate the likelihood (\autoref{eq:heatstress-likelihood}).
    % TODO add "Density" to y label in C
    }

     \label{fig:seeds}
 \end{figure}

>>>>

% TODO cite https://www.biorxiv.org/content/10.1101/2020.10.07.327759v1
% TODO cite Yang, F. et al. The fitness costs and benefits of trisomy of each Candida albicans chromosome. Genetics 218, 1–7 (2021). https://academic.oup.com/genetics/article/218/2/iyab056/6218773

@@

%%%%%%%% Fig illustation
\begin{figure}[b!]
  %\centering
  \begin{subfigure}[t]{0.4\textwidth}
      \caption{
        \textbf{A. Single-locus model}
      }
      \centering
      \includegraphics[height=1.8in]{figures/Fig1-A.pdf}      
      \label{fig:model1}
  \end{subfigure}
  %\\
  \begin{subfigure}[t]{0.4\textwidth}
  	  \caption{
        \textbf{B. Multi-locus model}
      }
      \centering
      \includegraphics[height=1.8in]{figures/Fig1-B.pdf}      
      \label{fig:model2}
  \end{subfigure}
  \caption{
    \textbf{Model illustrations.}
    \textbf{(A)} In the single-locus model, the four genotypes are: euploid wild-type, \euwt; euploid mutant, \eumt; aneuploid wild-type, \anwt; and aneuploid mutant, \anmt.
    Overall there are two possible trajectories from \euwt\; to \eumt.
    \textbf{(B)} In the multi-locus model, each genotype is characterized by its karyotype, \euwt\; or \anwt, and the number of accumulated beneficial mutations, denoted by stars. 
    In both panels arrows denote transitions between genotypes, with transitions rates: $\mu$, beneficial mutation rate; $\delta$, aneuploidy rate.
  }
  \label{fig:models}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Prior distributions.} We used informative prior distributions for $w_{2n+1}=1-c+b$, $w_{2n+1^*}=(1+s)(1-c)+b$ and $w_{2n^*}=1+s$, which we estimated from growth curves data from mono-culture growth experiments previously reported by \citet[Figs. 3C, 4A, and S2]{Yona2012}.
We used \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. Briefly, \texttt{Curveball} takes growth curves of two strains growing separately in mono-culture and predicts how they would grow in a mixed culture, that is, it predicts the results of a competition assay.
From these predictions, relative fitness values can be computed. Because \texttt{Curveball} uses a maximum-likelihood approach to estimate model parameters, we were able to estimate a distribution of relative fitness values by sampling from a truncated multivariate normal distribution defined by the maximum-likelihood covariance matrix. We sampled 10,000 samples to use as a prior distribution~(\autoref{fig:growth-curves}).
We used growth curves of \euwt\ and \anwt\ in \SI{39}{\celsius} to estimate a prior distribution for $w_{2n+1}$. In lieu of a better prior, we used the same prior for $w_{2n+1^*}$ and $w_{2n^*}$. 
To increase computational efficiency, we also assumed $w(2n*)>w(2n+1*)>w(2n+1)>w(2n)$; running the inference without this assumption produced similar results. % TODO IK is this correct?

@@

% FIG posterior
\begin{figure}[h!]
  \centering
\includegraphics[width=0.9\textwidth]{figures/posterior.pdf}
  \caption{
  \textbf{Posterior distribution of single-locus model.}
% TODO histogram on diagonal, 2d histogram without interpolation below diagonal
% TODO mutation rate tick labels should be in 10^-6, not 10^-5, and write 10^-6 not 1e-5; also it looks like you get positive density for mutation rate = 0 which doesnt make sense - set limit to 0.
% change trisomy rate to aneuploidy gain rate
% remove "mutation rate" from y label on top-left
% remove the y-ticks from the diagonal, they are meaningless, right?
% Fitness of 2n+1, w_{2n+1} (not 2n+1 fitness)
% TODO change dashed orange to red and make it a bit bolder.
On the diagonal, the posterior kernel density estimate for each parameter. 
Below the diagonal, joint posterior density of two parameters (dark purple and bright yellow for low and high density, respectively). Red markers and orange lines for the joint MAP estimates (which may differ from the marginal maximum density, as the marginal distribution integrates over all other parameters).
} 
  \label{fig:posterior}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%

@@

However, a model without aneuploidy where the aneuploidy rate is fixed at zero, $\delta=0$, cannot explain the experimental observations~(Figure~\ref{fig:2n*-fixation}). The estimated parameter values without aneuploidy are: 
% generated by MAP-and-hdi.ipynb
mutation rate, $\mu=7.837\substack{+0.078 \\ -0.123}\times10^{-9}$,
fitness of mutant, $w_{2n^*}=1.013\substack{+0.000 \\ -0.000}$. 
We can see that the mutation rate estimate in this model is much lower than in the model with aneuploidy. Higher mutation rate would cause quicker fixation of \eumt\ than in the experimental observations (\autoref{fig:2n*-fixation}). Thus, even disregarding the appearance of trisomic cells in the experiment, the evidence supports the inclusion of aneuploidy in the model.
% TODO WAIC?

% Fig single-locus model with and without aneuploidy
%%% generated by with-aneuploidy_vs_no_aneuploidy.ipynb

\begin{figure}[h!]

  \begin{subfigure}{0.5\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/fixation-plot-a.pdf}      
      \label{fig:fit}
  \end{subfigure}
  \begin{subfigure}{0.5\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/fixation-plot-b.pdf}      
      \label{fig:fit-cumulative}
  \end{subfigure}
  \caption{
    \textbf{Single-locus model fit with and without aneuploidy.}
    \textbf{(A)} The distribution of time to fixation of \eumt\ (i.e., adaptation time) in $10,000$ simulations of the single-locus model with aneuploidy (blue; MAP parameters) compared to two models without aneuploidy: a model with the same parameter values except $\delta=0$ (orange), and a model fitted to the data assuming $\delta=0$ (green).
In the experiment by \citet{Yona2012}, one population lost aneuploidy by generation 1,700 and another by generation 2,350 (dashed lines) but not before generation 450. Thus, the blue distribution is a better fit compared to the green, and the yellow histogram has a very poor fit.
The last bin contains all the simulations with time equal or greater than 3,000.
    \textbf{(B)}
	Cumulative distribution of the time to fixation of \eumt\ in 10,000 simulations using the MAP estimate with and without aneuploidy in blue and green, respectively, and corresponding to the blue and green bars in panel A. The MAP likelihood (\autoref{eq:heatstress-likelihood}) is 0.84 and 0.67 with and without aneuploidy, respectively.
     }
  \label{fig:2n*-fixation}
\end{figure}

\begin{figure}[h!]
  \centering
\includegraphics[width=0.8\textwidth]{figures/dynamics.pdf}
  \caption{
  \textbf{Single-locus model posterior genotype frequency dynamics.} 
  The posterior predicted frequency of the four genotypes. Each of 20 curves is the average of 10,000 simulations using parameters drawn from the posterior distribution (\autoref{fig:posterior}).
  }
  \label{fig:ppc-plot}
  % TODO put legend inside figure - I think there is room in middle right
\end{figure}

\paragraph*{Analysis: single-locus model.}
Here, we aim to analyse the timing of the single-locus model and the relative importance of the various parameters. As a stochastic model, there are two events that can occur from the wild type population. One is mutation that gives the adapted population state, or a missegregation event. The probability with which these events happen is hugely different. We are interested in the average waiting times until one of the two events occurs and lead to fixation. Let $T_1$ and $T_2$ denote the random variables for the waiting time for the events mutation and missegregation respectively. Due to discrete generation, they are geometrically distributed with parameter compound of the probability that at least one individual in this generation changes its state and goes to fixation:  
\begin{subequations}
\begin{align}
T_{2n^*}&\sim \text{Geo}(1-(1-2\mu \Delta w)^N)=\text{Geo}(p_{2n^*})\\
T_{2n+1}&\sim \text{Geo}(1-(1-2\delta\Delta w)^N)=\text{Geo}(p_{2n+1})
\end{align}
\end{subequations}
Here, $\Delta w$ denotes the relative difference between the fitness of the newly introduced type and the population mean fitness.
However, since we have competing events, we need to find the minimum of these two processes. The minimum of two geometrically distributed processes with parameters $p_{2n^*}$ and $p_{2n+1}$ is again geometrically distributed with parameter $1-(1-p_{2n^*})(1-p_{2n+1})$. The average waiting time until the first event happens is the inverse of the paramter of the geometric distribution. The probability that aneuploidy will be the first event to happen, is given by
\begin{equation}
p_A=\frac{p_{2n+1}}{p_{2n^*}+p_{2n+1}}=\frac{1-(1-2(w_3-w_1)\delta)^N}{2-(1-2(w_3-w_1)\delta)^N-(1-2(w_2-w_1)\mu)^N}.
\end{equation}
Since we are interested in the average time to fixation of $2n^*$, the steps that are needed differ depending on the actual event that happens first. If mutation happens first, then $2n^*$ is already reached and everything is fine. On the other hand, if aneuploidy occurs first, the population is in state $2n+1$ and several other steps have to occur, until $2n^*$ is reached. Thus, in this aneuploid trajectory, we also need to add the average waiting times to reach $2n+1^*$ and finally $2n^*$ The formulas are written down straightforward.

In general, we assume, that the waiting time until a new type occurs is much longer than the time the population needs to go to fixation for a certain type, but nonetheless, these times need to be included for a rigorous analysis. There are two ways, how to compute them and they differ in the assumption of the underlying process.

 If the population evolves deterministically, then one can derive the formula
\begin{equation}
t_\text{fix, i}=2\frac{\log(N)}{\log(1+s)}
\end{equation}
 from the classical one-locus two-allele equations.  As we work with simulations, we assume fixation if it reached 95\% of the total population size. This is significantly different from full fixation. Thus we need a slightely different formula
 \begin{equation}
 t_\text{fix, $f_c$, i}=\frac{\log\left(\frac{f_cN}{1-f_c}\right)}{\log(1+s)}.
 \end{equation}
 Here, $f_c$ denotes the fraction for which we assume fixation.

In the second approach, the population is assumed to evolve stochastically. It is based on the diffusion equation and given in Kimura and Ohta 1969 (eq. 17). We denote it by $T_\text{fix, i}$.

Thus, the average time to $2n^*$ over the aneuploidy trajectory is given by
\begin{subequations}\label{T_aneu_mut}
\begin{equation}
T_\text{aneu}=(1-(1-p_{2n^*})(1-p_{2n+1}))^{-1}+p_{2n+1^*}^{-1}+p_{2n^*}^{-1}+T_\text{fix, 2n+1}+T_\text{fix, $2n+1^*$}+T_\text{fix, $2n^*$}.
\end{equation}
The average time to $2n^*$ over the mutant trajectory is given by 
\begin{equation}
T_\text{mut}=(1-(1-p_{2n^*})(1-p_{2n+1}))^{-1}+T_\text{fix, $2n^*$}.
\end{equation}
\end{subequations}
\textcolor{red}{[MP: $T_\text{fix, $2n^*$}$ is not the same in (11) and (12). Need a better notation.]}

In summary, the average time to $2n^*$ is given by
\begin{equation}\label{T_total}
T=p_AT_\text{aneu}+(1-p_A)T_\text{mut}.
\end{equation}

\begin{remark} There are several implicit assumptions that the equations hold like this.
\begin{enumerate}
\item The two processes fixation and occurrence of a new type are sequential and do not interfere with each other.
\item There is no clonal interference.
\item The fixation process starts from exactly one individual. 
\end{enumerate}
There are certainly combinations of the mutation rate, the missegregation rate, the selection strength and the population size for which at least one of these assumptions is violated. Thus, \eqref{T_aneu_mut} and \eqref{T_total} may not yield a good approximation to the actual time to the adapted state.
\end{remark}

\begin{figure}[h!]
\centering
\includegraphics[width=.49\textwidth]{../Simulations/figinvpNonlymut0.pdf}
\includegraphics[width=.49\textwidth]{../Simulations/figinvpmuonlymut0.pdf}
\caption{The average time to $2n^*$ is shown as a function of the population size N in the left panel and as a function of the mutation rate in the right panel. The blue circles are the means of the simulation runs and the lines denote the span between the 5th and the 95th percentile. The orange line is the respective analytic approximation. In these two plots, the aneuploidy rate is zero, so only the mutational trajectory is available.}
\label{fig:only mut}
\end{figure}

To get a feeling of the limitations of \eqref{T_total}, we plot simulations with $\delta=0$ in Fig.~\ref{fig:only mut}. The other paramters are similar to those infered above. Even there, thereare slight deviations for high paramter values. Clearly visible in the left panel and not so good to see in the right. We conjecture, that in this parameter range more than one individual mutates in one generation and our assumption for the fixation time, which starts from exactly one mutant, does not hold.

@@

\begin{figure}[h]
% Fig Curveball 30deg
% generated with growth_curves.ipynb
\begin{subfigure}{0.5\textwidth}
    \centering
	\includegraphics[width=\textwidth]{figures/evo39_fitness_30deg.pdf}
    \label{fig:growth-curves-30deg}
\end{subfigure}
% Fig Curveball 39deg
% generated with growth_curves.ipynb
\begin{subfigure}{0.5\textwidth}
    \centering
	\includegraphics[width=\textwidth]{figures/evo39_fitness_39deg.pdf}
    \label{fig:growth-curves-39deg}
\end{subfigure}
\caption{
    \textbf{Fitness estimation from growth curves.}
    \textbf{(A)} Fitness estimation from growth curves collected at \SI{30}.{\celsius}
    \textbf{(B)} Fitness estimation from growth curves collected at \SI{39}.{\celsius}
    Growth curves previously described in \citet[Figs. 3C, 4A, and S2]{Yona2012}.
	Fitness estimated from growth curves using \texttt{Curveball}, a method for predicting results of competition experiments from growth curve data~\citep[\href{https://curveball.yoavram.com}{curveball.yoavram.com}]{Ram2019}. See \emph{Models and Methods, Prior distributions} for more details.
} 
\label{fig:growth-curves}
\end{figure}
%TODO please report here the point estimate and 95% CI for the three distributions, I only have for 1-c+b: 0.9863 (95% CI 0.9790 - 0.9939).
% TODO add panel labels

% Fig likelihood profile
\begin{figure}[p]
  \centering
  \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/sensitivity-A.pdf}      
      \label{fig:sensitivity-mutation}
  \end{subfigure}
  \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/sensitivity-B.pdf}      
      \label{fig:sensitivity-aneuploidy}
  \end{subfigure}
  \\
   \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/sensitivity-C.pdf}      
      \label{fig:sensitivity-anwt}
  \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/sensitivity-D.pdf}      
      \label{fig:sensitivity-anmt}
  \end{subfigure}
    \begin{subfigure}{0.3\textwidth}
      \includegraphics[width=\textwidth]{figures/sensitivity-E.pdf}      
      \label{fig:sensitivity-eumt}
  \end{subfigure}
  \caption{
    \textbf{Likelihood profiles.} Sensitivity of the approximate likelihood, $\likelihood(\theta)$, of the single-locus model to changing a single parameter while the other parameters remain fixed at their MAP estimates. Dashed vertical line represents the MAP value.
  }
  % TODO add 10^-10, not 10^-2
  % TODO ylabel - L shouldn't have a subscript !, this was used only for the model without aneuploidy
  % TODO add more points in the peaks

  \label{fig:sensitivity}
\end{figure}

% Fig ABC iterations, convergence
%% generated by convergence.ipynb
\begin{figure}[h]

  \begin{subfigure}{1\textwidth}
    \centering
      \includegraphics[width=0.45\textwidth]{figures/convergence-p1_mr.pdf}      
      \includegraphics[width=0.45\textwidth]{figures/convergence-p2_tr.pdf} \\     
      \includegraphics[width=0.325\textwidth]{figures/convergence-p3_w1.pdf}      
      \includegraphics[width=0.325\textwidth]{figures/convergence-p4_w2.pdf}      
      \includegraphics[width=0.325\textwidth]{figures/convergence-p5_w3.pdf}      
     \label{fig:convergence-A}
    \end{subfigure}
    \\    
  \begin{subfigure}{1\textwidth}
  	\includegraphics[width=\textwidth]{figures/ess.pdf}
     \label{fig:convergence-B}
  \end{subfigure}
    \caption{
    \textbf{Single-locus model inference convergence.} 
    The ABC-SMC algorithm was used to infer the parameters of the single-locus model. \textbf{(A-E)} The approximate posterior distributions of model parameters at each iteration of the ABC-SMC algorithm demonstrates convergence, as the posterior did not significantly change after the first iteration, $t=1$.
    \textbf{(F-I)} ABC-SMC measures of convergence. After iteration number 6, the acceptance threshold was $\epsilon=0.13$, the acceptance rate was $0.018$, the number of particles was 982, and the effective sample size ESS=651.
}
    \label{fig:convergence}
    % TODO legend - t=6, etc (add =)
	% x labels missing for G H I - either add them or put "Iteration" at the middle of the axes. whatever is easier.
\end{figure}

% Fig Validation of posteriors
\begin{figure}[h]
    \centering
	\includegraphics[width=0.6\textwidth]{figures/seeds.pdf}
    \caption{
    \textbf{Posterior distribution validation.}
    The posterior distributions of the single-locus model parameters are roughly the same regardless of the number of simulations (10-1,000) used to approximate the likelihood.
    } 
    \label{fig:seeds}
\end{figure}
% TODO put legend in panel A top right corner
% panel labels A->E
% y label "Density"; y ticks are meaningless, right? 
% x labels - see comments above. add symbols, fitness of 2n*, etc.
% color by # of reps; second run with 1000 reps same color as 1st but with dashed line
